---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, setup, include = FALSE}

knitr::opts_knit$set(root.dir = 'assessment')

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%")
```

# jjmR

```{r}
getwd()
```

<!-- badges: start -->
<!-- badges: end -->

The goal of jjmR is to ... Graphics and diagnostics libraries for SPRFMO's JJM model adopted from IMARPE's jjmTools


## Installation

You can install `jjmR` like so:

``` r
install.packages("devtools")
devtools::install_github("SPRFMO/jjmR")
```

## Basic Usage


`jjmR` provides support for objects of class `jm.output`. Reading in the outputs of a JJM model with `jjmR` provides a number of summary and plotting functions. 

In order to run any of the `jjmR` functions for reading assessment results, you need to make sure your working directory is set to the location of the `assessment` folder. 

```{r, eval=TRUE}
library(here)
library(jjmR)
# setwd(here("assessment"))


jjm_results <- readJJM("h2_1.13", path = "config", input = "input")

summary(jjm_results)
plot(jjm_results)

```


## Example of tidying jjmR outputs


`jjmR` also includes a series of helper functions to "tidy" up the data (see https://r4ds.had.co.nz/tidy-data.html). These are designed to help users access and comapre results across multiple models and stocks. 

First, let's load up two different model runs. As always with `jjmR`, make sure your working directory is set to the `assessment` folder. 


```{r, eval=TRUE}
library(jjmR)
library(ggplot2)
library(dplyr)
library(tidyr)
theme_set(theme_jjm(base_size = 15))
# example of code to generate model results
# model_results_1 <- runit(mod="h1_1.13",pdf=TRUE,est=TRUE,exec="../src/jjms")

# model_results_2 <- runit(mod="h2_1.13",pdf=TRUE,est=TRUE,exec="../src/jjms")

# read in a model run

model_results_1<- readJJM("h1_1.13", path = "config", input = "input")

model_results_2 <- readJJM("h2_1.13", path = "config", input = "input")

# compare to another model run

m1_v_m2 <- combineModels(model_results_1,model_results_2)
```


We can now use the `tidy_JJM` function to tidy up the results. As of now `tidy_JJM` provides support for a number of the most commonly used outputs of the JJM model, but not all. 

`tidy_JJM` returns a list with objects for each tidied data type (e.g. selectivities and msy_mt), with columns denoting the model, stock, etc. for a given observation. 

```{r}

tidy_jjm <- tidy_JJM(m1_v_m2)
str(tidy_jjm)

```


This tidy form is useful for users wishing to quickly generate new plots or analyses based on the results of different model runs. 

For example, suppose we want to plot the predicted index values for each of the indicies across models, as well as the residuals. 

```{r}
index_fits <- tidy_jjm$index_fits

index_fits %>% 
  ggplot() + 
  geom_pointrange(aes(year, observed_ind, ymin = observed_ind - 1.96 * observed_se, ymax =  observed_ind + 1.96 * observed_se)) +
  geom_path(aes(year, pred_ind, color = model)) + 
  facet_wrap(~ fleet_name, scales = "free_y") + 
  scale_x_continuous(name = "Year", guide = guide_axis(n.dodge = 2)) + 
  scale_y_continuous(name = "Index Values")

index_fits %>% 
  mutate(residual = pred_ind - observed_ind ) %>% 
  group_by(fleet_name, model) %>% 
  mutate(standardized_residual = residual / sd(residual)) %>% 
  filter(!is.na(standardized_residual)) %>% 
  ggplot() + 
  geom_hline(yintercept = 0,linetype = 2) +
  geom_col(aes(x = year, y =standardized_residual, fill =model), position = position_dodge(width = 0.5)) +
  facet_wrap(~ fleet_name, scales = "free_x") + 
  scale_x_continuous(name = "Year", guide = guide_axis(n.dodge = 2)) + 
  scale_y_continuous(name = "Standardized Residuals")

```

### Using `get_` Functions: Plotting Selectivities

You can also access the helper functions the underly `tidy_JJM` directly. Each of these is named `get_X`, where `X` is the name of the data. 

For example, to access the estimated selectivity ogives ,we can use `get_selectivities`. 

```{r}


selectivities <- get_selectivities(m1_v_m2)

head(selectivities)

```


Selectivities alone have a dedicated plotting function added to them by the "tidy" functions of `jjmR`. You can use this by running `plot_selectivities`, or by calling `plot` directly. 



```{r}

plot_selectivities(selectivities)

plot(m1_v_m2,what="selectivity",fleet="fsh", alpha = 0.2, scale = 10,
     years = 2000:2020)


plot(model_results_2,what="selectivity",fleet="ind", alpha = 0.2, scale = 10,
     years = 2015:2020)

```


### Miscelanneous Examples

```{r}
msy_my_results = get_msy_mt(m1_v_m2)

head(msy_my_results)

kobe(model_results_1, engine = "ggplot")

qs = get_catchabilities(model_results_1)

qs %>% 
  ggplot(aes(year, q, color = model)) + 
  geom_line() + 
  facet_wrap(~ fleet_name, scales = "free_y")


totals <- get_totals(m1_v_m2)

totals %>% 
  ggplot(aes(year, value, color = stock, linetype = model)) + 
  geom_line() + 
  facet_wrap(~ metric, scales = "free_y")


index_fits <- get_index_fits(m1_v_m2)

index_fits %>% 
  ggplot() + 
  geom_pointrange(aes(year, observed_ind, ymin = observed_ind - 1.96 * observed_se, ymax =  observed_ind + 1.96 * observed_se)) +
  geom_path(aes(year, pred_ind, color = model)) + 
  facet_wrap(~ fleet_name, scales = "free_y")


age_fits <- get_age_fits(m1_v_m2)


age_fits %>% 
  filter(model == "h2_1.13", stock == "Stock_1", year > 2015) %>% 
  pivot_longer(predicted:observed) %>% 
  ggplot() + 
  geom_density(aes(age, value, fill = name),stat = "identity", alpha = 0.5) + 
  facet_grid(year~fleet_name)


recruits <- get_recruits(m1_v_m2)

recruits %>% 
  ggplot() + 
  geom_ribbon(aes(year, ymin = lower_recruits, ymax = upper_recruits, fill = stock),alpha = 0.5) + 
  geom_line(aes(year, recruits, color = stock)) + 
  facet_wrap(~model)



fishing_mortality <- get_fishing_mortality(m1_v_m2)

fishing_mortality %>% 
  ggplot(aes(year, mortality, color = age, group = age)) + 
  geom_line() + 
  facet_grid(model~stock)


```



