---
title: "jjmR Code Review"
format: html
execute:
  eval: false
---

## Scope

Review of the R package source in `jjmr/`, focusing on likely bugs, robustness, and maintainability issues. Notes below reference file names and functions.

## Executive summary

Primary risks are hard failures during routine use: (1) `readJJMConfig.jjm.output()` errors immediately due to an undefined `output`; (2) `.runJJM()` can build an empty command when `adflags = NULL`; and (3) `readJJM()` can fail before reading outputs because `normalizePath()` expects existing files. Secondary issues affect robustness and reproducibility (assumptions about fixed column counts and shared fleet metadata), plus maintainability items (use of `eval(parse())`, strict class checks). Fixes are small, localized, and low‑risk.

## Summary (high‑priority items)

- **Confirmed runtime error** in `readJJMConfig.jjm.output()` due to an undefined `output` variable.
- **Likely runtime error** in `.runJJM()` when `adflags = NULL` (default): the current check uses `exists("adflags")`, which is always `TRUE` and can collapse the command to `character(0)`.
- **Potential failure** in `readJJM()`/`.getOutFile()` when output files are missing because `normalizePath()` defaults to `mustWork = TRUE`.

## Bugs and fixes

### 1) `readJJMConfig.jjm.output()` references `output` but never defines it

**File:** `R/readJJMConfig.R`

```{r}
readJJMConfig.jjm.output = function(model, path, input=NULL, ...) {
  ctl  = .getCtlFile(model=model, path=path)
  dat  = .getDatFile(ctl=ctl, input=input)
  
  reps = .getRepFiles(model=model, output=output) # output is undefined
  yld  = .getYldFile(model=model, output=output)
  ...
}
```

**Issue:** `output` is not in the argument list for this method, so the function errors immediately.

**Fix:** add `output = "results"` to the signature and pass it through, or use `output` from the model object (if stored there).

---

### 2) `.runJJM()` command assembly breaks when `adflags = NULL`

**File:** `R/jjmr-internal.R` (`.runJJM`)

```{r}
if(exists("adflags")) {
  jjm = paste0(jjm, " ", adflags)
}
```

**Issue:** `exists("adflags")` is always `TRUE` inside the function. When `adflags` is `NULL` (default), `paste0(jjm, " ", NULL)` returns `character(0)`, which breaks the `system()` call.

**Fix:**

```{r}
if (!is.null(adflags)) {
  jjm <- paste0(jjm, " ", adflags)
}
```

---

### 3) `readJJM()` fails early if outputs do not exist

**File:** `R/jjmr-internal.R` (`.getOutFile`)

```{r}
.getOutFile = function(model, output, ext) {
  out = file.path(output, paste0(model, ext))
  out = normalizePath(out)  # mustWork = TRUE by default
  return(out)
}
```

**Issue:** `normalizePath()` errors when the file does not exist. This causes `readJJM()` to error before a clearer diagnostic can be given.

**Fix:** Use `normalizePath(..., mustWork = FALSE)` and follow with an explicit `file.exists()` check with a targeted error message.

---

### 4) Unused `version` argument in `readJJM()`

**File:** `R/readJJM.R`

`readJJM()` accepts `version`, but it always calls `.versionJJM(ctl)` and ignores the user-supplied value.

**Fix:** Either remove the argument or respect it (e.g., `version = version %||% .versionJJM(ctl)`), and document precedence clearly.

---

### 5) `compareModels()` uses `eval(parse())`

**File:** `R/compareModels.R`

```{r}
cmd <- paste0("mods_comb <- combineModels(",
              paste(paste0("temp[[",seq_along(mods),"]]"),collapse=", "),")")

eval(parse(text=cmd))
```

**Issue:** `eval(parse())` is avoidable and makes debugging harder.

**Fix:**

```{r}
mods_comb <- do.call(combineModels, temp)
```

---

### 6) `get_index_fits()` assumes 6 V-columns without checking

**File:** `R/get_index_fits.R`

```{r}
names(things)[grepl("^V",names(things))] <- c('year', "observed_ind", "pred_ind", "observed_se", "pearson_resid", "something_resids")
```

**Issue:** If the number of `V` columns changes, this silently mislabels columns or errors.

**Fix:** check `sum(grepl("^V", names(things))) == 6` and stop with a clear message if not, or dynamically map by position with validation.

---

### 7) Class checking in `.combineModels()` is too strict

**File:** `R/jjm.output-internal.R`

```{r}
for(i in modelList)
  if(class(i) != "jjm.output")
    stop("Objects must be of class 'jjm.output'.")
```

**Issue:** `class(i)` may include multiple classes. `inherits(i, "jjm.output")` is safer.

**Fix:**

```{r}
if (!inherits(i, "jjm.output")) stop("Objects must be of class 'jjm.output'.")
```

---

### 8) Fleet name mapping assumes all models share identical fleet metadata

**Files:** `R/get_age_fits.R`, `R/get_index_fits.R`, `R/get_len_fits.R`, `R/get_selectivities.R`

**Issue:** Fleet name lookup uses `models[[1]]$data$Inames/Fnames` for all models. If models differ (common in sensitivity runs), joins become incorrect or incomplete.

**Fix:** Build `fleet_names` per model inside `top_getter()` and join before concatenating across models, or validate that all models share identical fleet definitions and throw a clear error if not.

## Fix roadmap (ordered)

1. Fix `readJJMConfig.jjm.output()` signature and output handling.
2. Correct `.runJJM()` adflags check to avoid empty command strings.
3. Make `.getOutFile()` tolerant of missing outputs and add explicit existence checks in `readJJM()`.
4. Replace `eval(parse())` in `compareModels()` with `do.call()`.
5. Add input validation in `get_index_fits()` for expected column counts.
6. Use `inherits()` in `.combineModels()` class checks.

## Improvements (non‑breaking)

- **Explicit returns:** Many functions end with an assignment (e.g., `out <- ...`). This works but is fragile. Prefer `return(out)` for clarity.
- **Input validation:** Add checks for missing/empty output files with actionable messages in `readJJM()`, `readJJMConfig()`, and `runJJM()`.
- **Consistent parameter handling:** A few functions accept `version` but do not use it. Either remove or standardize across all functions.
- **Avoid `setwd()` side effects:** `.runJJM()` changes directories without restoring. In `runJJM.default` the caller resets working directory, but adding `on.exit(setwd(oldwd))` within `.runJJM()` would make the function safer on its own.
- **Testing:** Add minimal `testthat` tests for:
  - `readJJMConfig.jjm.output()` works with a small fixture.
  - `runJJM()` assembles commands correctly when `adflags = NULL` and when set.
  - `get_index_fits()` fails cleanly when the expected columns are not present.

## Suggested quick patch list

1. Fix `readJJMConfig.jjm.output()` signature and `output` usage.
2. Replace `exists("adflags")` with `!is.null(adflags)` in `.runJJM()`.
3. Update `.getOutFile()` to `normalizePath(..., mustWork = FALSE)` and add explicit `file.exists()` checks.
4. Replace `eval(parse())` in `compareModels()` with `do.call()`.
5. Add column count checks in `get_index_fits()` (and other `get_*` functions that assume fixed column counts).
6. Use `inherits()` in `.combineModels()` class checks.

